Sperm should evolve to make female meiosis fair.

### 2014-11-17 writeup_and_notes/evolution/response_drive_MS.tex

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\bibliographystyle{evolution}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
	\newcommand{\gc}[1]{{ \color{red} #1}}
	\newcommand{\yb}[1]{{ \color{blue} #1}}
\usepackage{lineno}
%\linenumbers 
\begin{document}
Dear EVOLUTION editorial team,
\\

Thank you for your consideration of our manuscript 14-0566 entitled "Sperm should evolve to make female meiosis fair." We appreciate the thorough comments by the reviewers and Associate Editor.
We are submitting a new version of this manuscript in which we have incorporated reviewer comments as advised by the AE. Our revision includes versions of Models 1 and 2 with some degree of inbreeding, and has been edited to ensure clarity, and minimize the possibility of misleading our readers. We believe that this has resulted in a stronger manuscript. Below we describe in greater detail, and a point-by-point basis how we responded to specific comments of the AE and the reviewers.
\\
\\
Sincerely \\
Yaniv Brandvain and Graham Coop
\newpage

{\bf{Directive from Editor}} \\ \\
Montpellier, Nov. 6th 2014,
\\ \\
Dear Mr. Brandvain: 
I am pleased to tell you that your manuscript 14-0566 entitled "Sperm should evolve to make female meiosis fair." will be suitable for publication in Evolution pending some additional, minor revision. We appreciate your submission as well as your willingness to respond to the constructive suggestions of the reviewers and Associate Editor, Dr. Jon Wilkins.

You will find the comments from the Associate Editor and the Reviewers below. I invite you to respond to these and revise your manuscript accordingly. If your responses require additions to the text of the manuscript, please take care not to exceed our manuscript length guidelines, as these apply to the final, accepted version of the manuscript. \dots{}

Once again, thank you for submitting your manuscript to Evolution.  I look forward to the published paper.
\\  \\
Sincerely, \\
Dr. Thomas Lenormand \\
Editor, Evolution \\
thomas.lenormand@cefe.cnrs.fr

\newpage

{\bf{Comments from the Associate Editor}} -- \emph{Authors respond in italics}  \\

Thanks for the interesting paper. As both reviewers note, it is an interesting and important topic. I also found the biological framing compelling, and I think it should penetrate deeper into the non-theory part of the Evolution audience than many theory papers do.

Let me apologize for the slowness of the review process here. I had a heck of a time getting reviewers to agree to this one. I don't know. Maybe they somehow saw that the combined PDF was 97 pages and got scared off. Or maybe when they read the abstract and saw the umlauts on oogenesis they assumed that this was actually a mis-addressed New Yorker submission.
\\ \\
\emph{We appreciate your feedback and efforts to co\"{o}rdinate reviewers}
\\

You've got some nice reviewer comments here, which should definitely be addressed. In particular, reviewer 2, aka Gardner, made two suggestions that I see as critical.
\\ \\
{\bf{AE comment 1:}} First is the suggestion to consider inbreeding. Personally, I'm not sure it would be worth the effort at this point to write out and analyze a full-on model of a structured population with migration, etc, etc. It's a fantastic idea, and definitely worth doing, but maybe it's the next paper.

However, it would be easy to tweak each model with the inclusion of one or two F values, to give you excess homozygosity (or excess probability of green-beard matching). For example, in model 3, you mention this narrow region of parameter space that exists only when the paternal modifier functions similarly in sperm from heterozygous and homozygous fathers. One thing that inbreeding would do is increase the likelihood of occupying that space. Would that then make a sustained polymorphism biologically plausible? If so, what would that predict for the evolution of female meiosis in highly inbred populations?

{\bf{Response to AE comment 1:}}  \emph{We note that both R1 and R2 noted an interest in the case of inbreeding. \yb{We therefore have explored the behavior of model 2 [as well as model one for reference to the case of traditional drive] under six values of inbreeding. We briefly discuss these results in the modified text and make the figures [analogous to Figure 1B] in the Supplement.}}
\\
\\
\\
{\bf{AE comment 2:}} Second is the taxonomic question. The phylogenetic distribution of post-fertilization completion of female meiosis is critical to the evolutionary implications of this work. Even if the data is spotty at the moment, it would be worth spelling out what is known. Plus, you could maybe make your phylogenetic hypotheses a bit more specific that way.

{\bf{Response to AE comment 2:}}  \emph{WE NEED TO DO SOMETHING HERE \dots{}}
\\
\\
\\
{\bf{AE comment 3:}} 
A couple of comments regarding the discussion section:

In general, the discussion section is a little bit fast and loose with the evolutionary reasoning, and it is not always entirely clear what is being claimed. I get that you're trying to go for a readable voice here, to spell out the take-home message for the reader who only skimmed the results. However, I think this would benefit from thinking carefully about what the four or five key points are, and dedicating a paragraph to each.

{\bf{Response to AE comment 3:}}  \emph{We have included clear `sign-posts' throughout the discussion, and have more fully articulated the logic underlying our predictions and conclusions.}
\\
\\
\\
{\bf{AE comment 4:}} 
You mention at the end that you're not modeling female meiotic arrest. However, earlier in the discussion (ca. lines 309 \& 349) and at the end of the introduction (line 109), you make a couple of statements that border on the teleological. The current text borders on suggesting that delayed female meiosis could be favored because it permits the sperm to participate in the suppression of female meiotic drivers.

Now, that's a cool idea, but it suffers from the same limitation as evolvability arguments: it poses an adaptive benefit in terms of future selective events. It can work, but you have to be careful about the conditions. For example, either drivers would have to crop up often enough that the transient fitness reductions create a substantial fitness advantage -- or that there is a persistent fitness cost after fixation. And maybe it requires (or would benefit from) some higher-level selection.

Plus, you can flip the just-so argument around. We might expect that in species without post-fertilization completion of female meiosis, this is a consequence of female drivers that have accelerated meiosis in order to prevent interference by sperm modifiers.

Basically, there's a bunch of stuff here that would need to be explicitly modeled. So, just make sure all the appropriate caveats are in place.

{\bf{Response to AE comment 4:}}  \emph{We agree and have further clarified this point. This includes pointing out the circularity in logic of a simple version of the claim that female meiotic arrest evolves to allow sperm to combat female drivers. We do however, maintain our 3rd prediction [line 249 in our first submission] that reducing opportunities for drive could allow organisms to skip the fertilization requirement. We point out that this is a question of evolutionary maintenance rather than of evolution origin.}
\\
\\
\\
{\bf{AE comment 5:}}
More compelling in the context of the present work is the line of reasoning followed from line 354 onward, that the form of meiosis will affect the prevalence of female drivers of different sorts. I would actually like to see this expanded on a bit more.

The Om and In bit is nice, and a key piece of evidence supporting the applicability of your analysis. I do have a question regarding the predictions on crosses though. It seems that the prediction only holds to the extent that there are drivers that are in the process of being suppressed, which depends on the frequency with which new drivers arise and the rate at which they get suppressed (and presumably eventually eliminated). However, it seems like you might also expect to see the effect if there is some divergence between versions of the same drivers in different populations. Intuitively, if there is sperm-mediated suppression going on, there might be opportunities for antagonistic coevolution, with drivers selected to evade suppression (depending on the molecular mechanisms of drive and suppression, of course). This would make a similar set of predictions, I think, but would not hinge on a new driver having arisen in one of the two.

If the coevolution argument holds water, you would also expect to see faster evolutionary divergence of drive elements in species with post-fertilization meiosis. However, it might be hard to establish apples-to-apples comparisons to test that sort of thing.

{\bf{Response to AE comment 5:}} \emph{We agree with this ida, and have incorporated it into our discussion.}
\\
\\
\\
{\bf{AE minor comments:}}
\\
A couple of minor things
\\
throughout: hyphenate compound adjectives, e.g. "sperm-acting modifier" and "drive-suppressing allele"
\\
line 315: "outweighs"
\\
line 330: you forgot the gratuitous umlaut on "naive"
\\
line 372: extra space after )

{\bf{Response to AE minor comments:}} \emph{We fixed these, thanks!}

\newpage
Reviewer(s)' Comments to Author:
\\
Reviewer: 1
\\
Comments to the Author
\\Sperm favor fair female meiosis.

I enjoyed reading this interesting paper. The question is clearly one where intuition fails because of the complexity of the interacting factors and the authors? models clarify these interactions.
\\
\\
\\
{\bf{R1 comment 1:}}
Models 1-3 cover the situation in which drive in female meiosis depends on sperm (padumnal) genotype or male (paternal) genotype. The models find that the invasion of such a system of drive is unlikely because in a random-mating population drive is very rare because both sexual partners need to carry the rare allele and because drive, when it does occur, creates driving homozygotes which are assumed to have low viability and which cannot themselves undergo drive. If there is no cost to homozygotes the ?driver? should go to fixation slowly at first. Forms of inbreeding would probably also facilitate initial invasion.

{\bf{Response to R1 comment 1:}} \emph{We have added inbreeding to our analyses of models 1-2, see response to AE comment 1}
\\
\\
\\
{\bf{R1 comment 2:}}
Models 4-5 cover the situation in which an allele drives in female meiosis and this is modified by sperm genotype at the same locus (by assumption of tight linkage). Model 6 considers an unlinked sperm modifier. The interesting case in Model 4 is the suppressor allele B-. This allele initially does not suppress its own drive (because rare) but suppresses the drive of B and thus ends up in A/B- heterozygotes rather than B/B- heterozygotes. The authors say ``avoid the excess formation of low fitness homozygotes'' but I think this is slightly misleading because B and B- are different ?alleles?. B- displaces B and often comes to a lower equilibrium frequency presumably because of the effects of suppressing its own drive
(B- causes the production of higher fitness A/B- heterozygotes rather than lower fitness B-/B- homozygotes. If the difference in fitness is not too great then B- ?loses? from suppressing its own drive).

{\bf{Response to R1 comment 2:}} \emph{We have changed this section to (e.g. changing ``avoid the excess formation of low fitness homozygotes'' to ``avoid the excess formation of low fitness genotypes'') in accordance with these suggestions.}
\\
\\
\\
{\bf{R2 minor comments:}}
\\ line 272 ``so it avoid[s]'' line 274 ``the the''
\\
The take home message is that sperm alleles usually evolve to suppress female meiotic drive, whether linked (in coupling or repulsion) or unlinked to the driving locus.
\\
line 339 ``novel females meiotic drivers'' rather ``novel female meiotic drivers'' or ``novel females' meiotic drivers''
\\
line 342 ``Tthese''

{\bf{Response to R2 minor comments:}} \emph{We fixed these, thanks!}

\newpage{}

Reviewer: 2

Comments to the Author

{\bf{R2 summary:}}
This study raises an interesting question: are sperm favoured to meddle in female meiosis and, if so, what are they favoured to do? This question is motivated by the very interesting observation that, in many animals, female meiosis takes place after fertilization, giving the sperm some possible role in deciding the fate of meiotic drivers. The authors develop some mathematical models, considering cases in which sperm promote drive of matching alleles only or promote or suppress drive in general. The overall message coming from the models is that sperm do not act to promote drive, and can be favoured to suppress it.

Although I found the motivating question and background biology really interesting, I am not really convinced by the analysis.

{\bf{Response to R2 summary:}} \emph{We appreciate that you find the question, the biology, and the results compelling, and are disappointed that you are unconvinced by the analysis. However, we can find no statement in your specific comments below that suggest that you are unconvinced. We therefore respond to the comments below, and hope that the responses to your specific comments are convincing.}
\\
\\
\\
Some comments:
\\
{\bf{R2 comment 1 :}}
(1) The results are pretty obvious a priori, and don?t really warrant the development and analysis of explicit models. As the authors note in the Introduction, the matching-alleles models are instances of greenbeard genes, and we already know that these are under positively frequency-dependent selection, and cannot invade from rarity in unstructured populations if there is any obligate cost involved, such as the cost of the machinery that does the allelic recognition. (This link also needs to be made in the Results and Discussion sections.) And in the scenarios where sperm are deciding the fate of drivers in an egg to which they have no genetic relatedness, it is obvious that their only concern will be to maximize the viability of the ensuing zygote, which means suppressing drive.

{\bf{Response to R2 comment 1:}} \emph{We believe that intuition without rigorous theoretical support can be dangerous. Therefore, while the results may be intuitive to R2, they were not to ourselves or to R1. We therefore believe that our results are not trivial and represent an important contribution. The non-triviality of our results are apparent in this comment -- in our model drive presents no cost when extremely rear, and there is some amount of cost for which a sperm acting drive promoter can invade and fix. The complexity of our analysis [and the bit that can choke simple intuition] is the competition between two forms of frequency dependence in the costs and benefits of ``green-bearded drive''.}
\\
\\
\\
{\bf{R2 comment 2:}} (2) Explicit models would be a lot more useful if the scope of the study were extended to inbred populations, where there are more selective pressures at play and where it is less obvious to see what will be favoured. Here, driving genes would be a lot more likely to encounter identical-by-descent copies in sperm, which could act to promote collaboration on the part of the sperm (analogous to the boost that greenbeards get in structured populations, owing to increased rate of encounters). But drivers would also suffer from the reduced frequency of heterozygotes under inbreeding, and this would tend to inhibit drive. And there may be additional pressures owing to inbreeding favouring bias sex ratios, which driving sex chromosomes or, indeed, whole genomes could be important in bringing about, under female heterogamety e.g. see Gardner \& Ross (2014, ?Mating ecology explains patterns of genome elimination?, Ecol Lett, doi: 10.1111/ele.12383). An explicit model would be doing really useful work here, showing which of these effects is more important, and why. The present results for the outbreeding case, that seem a bit too trivial to stand on their own, would usefully serve to prepare the reader for more substantial results for inbreeding scenarios.

{\bf{Response to R2 comment 2:}} \emph{We agree that the case of inbreeding could be interesting. In line with comments from the AE and R1, we implement a simple model of inbreeding [see response to AE comment 1]; however, as stated by R1, completely constructing a detailed inbreeding model is beyond the scope of this ms and is best left to future work. }

\emph{As stated above, we disagree with the notion that a model which confirms the intuition of R2 is trivial. We believe that clearly and rigorously presenting novel predictions and highlighting relevant biology is a substantial contribution. We are also somewhat concerned that chasing special scenarios in the hopes of unintuitive results could district from our major and clear finding, and therefore limited our analysis of the case of inbreeding.}
\\
\\
\\
{\bf{R2 comment 3:}}  (3) I really enjoyed the biology discussed at the beginning of the Introduction. I think this could be fleshed out a lot more. For example, when you say ?in many animal species? (line 46), it would be useful to explain this more fully, perhaps with the help of a figure giving a taxonomic overview. Is post-fertilization female meiosis monophyletic, have there been reversals, are there obvious ecological or other correlates, and so on?

{\bf{Response to R2 comment 3:}}  \emph{\yb{NEED TO WORK ON THIS}}
\\
\\
\\
\\
As is my policy, I waive anonymity
\\
Andy Gardner

\end{document}

### 2014-11-17 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
added a new folder [evolution] with a draft of our evolution resubmision and response to reviewers

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\bibliographystyle{genetics}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
	\newcommand{\gc}[1]{{ \color{red} #1}}
	\newcommand{\yb}[1]{{ \color{blue} #1}}

@@

\newpage 
\flushleft{{\bf{Running title:}} Sperm prevent female meiotic drive} \\
{\bf{Keywords:}} Meiosis; Meiotic drive; Fertilization; Genomic conflict \\
{\bf{Corresponding author:}} Yaniv Brandvain. \emph{Address:} Department of Plant Biology, University of Minnesota --€“ Twin
Cities. St. Paul MN, 55108. \emph{Phone:} 612-624-4375 \emph{Email:} ybrandvain@gmail.com

@@

{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
O\"{o}genesis is inherently susceptible to such conflicts
because alleles compete for inclusion in the  product of female meiosis that will be transmitted to the egg. 
Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
In this light, many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
That females  of many animal species do not complete meiosis until after fertilization, 
     appears to run counter to this interpretation,
     because this delay provides an opportunity for sperm-acting alleles to
     meddle with the outcome of female meiosis and help like alleles drive in heterozygous females. 
Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
	in fact, sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. 
These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals.   
Rather than providing an opportunity for sperm collaboration in female meiotic drive,
        the `fertilization requirement'  indirectly protects females from
        meiotic drivers by providing sperm an opportunity to
        suppress drive. 
%	and suggest that the requirement of fertilization for the
%%        completion of female meiosis
%        potentially represents a mechanism employed by females 
%	 to ensure a fair meiosis.
\newpage

@@

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Current evidence from the best characterized systems of female meiotic drive in animals 
	(the  \emph{In} and \emph{Om} loci in mice) 
	suggests that sperm influence on female meiotic drive is not only possible, but likely. 
%Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
%There are current only two polymorphic, well-characterized female meiotic drive systems in
%animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
 While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages 52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
The fact that the two best characterized, polymorphic systems of putative female 
	meiotic drive systems in animals 
	show this effect suggests that if female meiotic drive is common 
	the role of sperm in modifying female drive will be important as well.  \newline

Numerous lines of evidence suggest that female meiotic drive is a common and important 
	evolutionary force, and therefore the opportunity for sperm to influence female drive is likely relevant to many animals.    
While research to date has identified a few extreme cases of female meiotic 
	drive in the small number model systems systematically
        studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
	rapid evolution of the basic components of the meiotic machinery (e.g. centromeres, telomeres, etc \dots) 
	suggest consistent selection on female meiotic drivers and suppressors of meiotic drive in many animal species 
	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}. 
 We expect that over the next decade the spread of sequencing to a
         range of systems will reveal many more female meiotic drive
         systems, however, carefully characterizing them will still
         remain a challenging task. 
\newline

%The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
%	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
%][]{Masui_book}). 
%It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
%In \emph{C. elegans}, premature deployment of the aster -- a vital component of mitotic
%machinery provided by the sperm, disrupts MII meiotic segregation
%in the egg, leading to a triploid zygote \citep{McNally2012}. 
%Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
% Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
% transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
% While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
%52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus, ][]{Pomiankowski1993}.
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg may function to enforce an equitable transfer of genetic material.

%%don't thi k this helped: 
%\yb{In fact, a mutant female which completes meiosis before 
 %	fertilization is giving up the protection from meiotic drivers potentially offered by sperm, 
%	and is therefore more likely to produce low fitness offspring. }
%%\gc{We do not claim that female meiotic arrest evolved
%  specifically to recruit helpful sperm to defend against meiotic
%  drive, the delayed completion of female meiotic drive will not in
%  fact encourage sperm to assist renegade drivers. Rather, we think
%  that features of
%  o\"{o}genesis  may evolve to facilitate an interaction with sperm 
%with female meiosis
%as yet another mechanism to enforce an equitable transfer of genetic material.}

@@

We present six related models -- three single-locus `pleiotropy' models and three two-locus `drive-modifier' models.
	{\bf{Model 1}} describes a single-locus female meiotic driver.  
	{\bf{Model 2}} describes a single-locus sperm-dependent female driver -- that is, an allele whose transmission in female meiosis depends on sperm haplotype. 
	{\bf{Model 3}} describes a single-locus paternal-dependent female driver -- an allele whose transmission in female meiosis depends on paternal genotype. 
Assuming that a traditional driver segregates at its equilibrium frequency (identified in Model 1), 
	we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) 
	sperm-dependent modifiers of drive. 
In {\bf{Models 4 and 5}}, we treat this two-locus system as if it consists of
	a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
	sperm-modifier is very tightly linked to the driving ({\bf{Model 4}}) or non-driving allele ({\bf{Model 5}})
 	at the drive locus such that recombination is unexpected. 
To evaluate the feasibility of sperm-modification of female meiotic drive, as compared to female suppression of drive, 
	we conclude with a model of female drive suppression by an unlinked female acting suppressor ({\bf{Model $6^\prime$}}). 
In all cases, we assume that fitness is independent of the drive modifier. 
 \newline

@@

We verbally describe our main results below.  
Readers interested in the details of these results should turn to the Appendix for a mathematical treatment, 
	and to our Mathematica worksheet  (File S2) for our complete derivations. 
There we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
Because a number of our analyses are approximations based on assuming
that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3.

@@

\begin{figure}
\begin{center}
\includegraphics[width = 1 \textwidth]{../Figures/selfprom.eps} \end{center}
\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
	Transmission probabilities for alleles through female
 	 meiosis depend on sperm genotype. 
	 The non-driving $A$-allele, and self-promoting $B$-allele
	 are represented by unfilled and filled circles, respectively. 
	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
	 	Assuming that the fitnesses of drive homozygotes and 
		heterozygotes are $1-s$ and $1$, respectively. 
	 	\emph{Main figure:} Boundary conditions for the invasion 
		and fixation of self-promoting and standard meiotic drivers, 
		with drive coefficient, $d$. Colored regions depict exact results, 
		while lines  represent analytical approximations. 
	 \emph{B1:}  Trajectories of sperm-dependent female drive
		 %alleles that can invade the population, 
		 each allele has $s =0.1$ against the
		 homozygotes.  The drive coefficient is denoted by color. 
 }   
\label{Eggsperm}
\end{figure}

@@

  \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}

>>>>

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\bibliographystyle{evolution}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
	\newcommand{\gc}[1]{{ \color{red} #1}}
	\newcommand{\yb}[1]{{ \color{blue} #1}}
\usepackage{lineno}
\linenumbers

@@

@@

{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
O\"{o}genesis is inherently susceptible to such conflicts
because alleles compete for inclusion into the egg. 
Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
In this light, many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
That females  of many animal species do not complete meiosis until after fertilization, 
     appears to run counter to this interpretation,
     because this delay provides an opportunity for sperm-acting alleles to
     meddle with the outcome of female meiosis and help like alleles drive in heterozygous females. 
Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
	in fact, sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. 
These results are consistent with the apparent sperm dependence of the best characterized female meiotic drivers in animals.   
Rather than providing an opportunity for sperm collaboration in female meiotic drive,
        the `fertilization requirement'  indirectly protects females from
        meiotic drivers by providing sperm an opportunity to
        suppress drive. 
\newpage

@@

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\yb{INSERT BIOLOGY HERE}
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Current evidence from the best characterized systems of female meiotic drive in animals 
	(the  \emph{In} and \emph{Om} loci in mice) 
	suggests that sperm influence on female meiotic drive is not only possible, but likely. 
 While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages 52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
The fact that the two best characterized, polymorphic systems of putative female 
	meiotic drive systems in animals 
	show this effect suggests that if female meiotic drive is common 
	the role of sperm in modifying female drive will be important as well.  \newline

Numerous lines of evidence suggest that female meiotic drive is a common and important 
	evolutionary force, and therefore the opportunity for sperm to influence female drive is likely relevant to many animals.    
While research to date has identified a few extreme cases of female meiotic 
	drive in the small number model systems systematically
        studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
	rapid evolution of the basic components of the meiotic machinery (e.g. centromeres, telomeres, etc \dots) 
	suggest consistent selection on female meiotic drivers and suppressors of meiotic drive in many animal species 
	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}. 
 We expect that over the next decade the spread of sequencing to a
         range of systems will reveal many more female meiotic drive
         systems; however, carefully characterizing them will still
         remain a challenging task. 
\newline

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus, ][]{Pomiankowski1993}.
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
\yb{Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg may result in an equitable transfer of genetic material -- 
	wether this result is the ultimate evolutionary function of the fertilization requirement or a coincidental pleiotropic outcome is beyond the scope of this manuscript, 
	but our intuition argues against the prior (see Discussion).}

@@

We present six related models -- three single-locus `pleiotropy' models and three two-locus `drive-modifier' models.
	{\bf{Model 1}} describes a single-locus female meiotic driver.  
	{\bf{Model 2}} describes a single-locus sperm-dependent female driver -- that is, an allele whose transmission in female meiosis depends on sperm haplotype. 
	{\bf{Model 3}} describes a single-locus paternal-dependent female driver -- an allele whose transmission in female meiosis depends on paternal genotype. 
Assuming that a traditional driver segregates at its equilibrium frequency (identified in Model 1), 
	we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) 
	sperm-dependent modifiers of drive. 
In {\bf{Models 4 and 5}}, we treat this two-locus system as if it consists of
	a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
	sperm-modifier is very tightly linked to the driving ({\bf{Model 4}}) or non-driving allele ({\bf{Model 5}})
 	at the drive locus such that recombination is unexpected. 
To evaluate the feasibility of sperm modification of female meiotic drive, as compared to female suppression of drive, 
	we conclude with a model of female drive suppression by an unlinked female-acting suppressor ({\bf{Model $6^\prime$}}). 
In all cases, we assume that fitness is independent of the drive modifier. 
 \newline

@@

We verbally describe our main results below.  
Readers interested in the details of these results should turn to the Appendix for a mathematical treatment, 
	and to our Mathematica worksheet  (File S2) for our complete derivations. 
There, we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
Because a number of our analyses are approximations based on assuming
that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximate. 
We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3.

@@

%%%FIG!

@@

\newpage
\section*{Figure Legends}
\clearpage
\begin{figure}
%\begin{center} \includegraphics[width = 1 \textwidth]{../Figures/selfprom.eps} \end{center}
\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
	Transmission probabilities for alleles through female
 	 meiosis depend on sperm genotype. 
	 The non-driving $A$-allele, and self-promoting $B$-allele
	 are represented by unfilled and filled circles, respectively. 
	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
	 	Assuming that the fitnesses of drive homozygotes and 
		heterozygotes are $1-s$ and $1$, respectively. 
	 	\emph{Main figure:} Boundary conditions for the invasion 
		and fixation of self-promoting and standard meiotic drivers, 
		with drive coefficient, $d$. Colored regions depict exact results, 
		while lines  represent analytical approximations. 
	 \emph{B1:}  Trajectories of sperm-dependent female drive
		 %alleles that can invade the population, 
		 each allele has $s =0.1$ against the
		 homozygotes.  The drive coefficient is denoted by color. 
 }   
\label{Eggsperm}
\end{figure}

\clearpage
\newpage
